Supporting information: Copulas and their potential for ecology

### 2019-09-19 SupportingInformation.Rmd
done a roudn of edits up through Concepts and Methods for Q1

<!--Now the copula figures that were not presented in the main text-->

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvClayton.pdf}
\caption[Example SC copulas]{Log-transformed pdfs (A-E) and samples (F-J) 
from example SC 
copulas. S is Spearman correlation; p is the value of the parameter for the SC 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively.\label{figped_SC}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Gumbel.pdf}
\caption[Example G copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example G copulas. S is 
Spearman correlation; p is the value of the parameter for the G family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_G}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvGumbel.pdf}
\caption[Example SG copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SG 
copulas. S is Spearman correlation; p is the value of the parameter for the SG 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively.\label{figped_SG}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Joe.pdf}
\caption[Example J copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example J copulas. K is Kendall correlation; p is the value of the parameter for the J family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_J}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvJoe.pdf}
\caption[Example SJ copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SJ 
copulas. K is Spearman correlation; p is the value of the parameter for the SJ 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively.\label{figped_SJ}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Frank.pdf}
\caption[Example F copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example F copulas. S is Kendall correlation; p is the value of the parameter for the F family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_F}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB1.pdf}
\caption[Example SBB1 copulas]{Log-transformed pdfs for example SBB1 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB6.pdf}
\caption[Example BB6 copulas]{Log-transformed pdfs for example BB6 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB6}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB6.pdf}
\caption[Example SBB6 copulas]{Log-transformed pdfs for example SBB6 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB6}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB7.pdf}
\caption[Example BB7 copulas]{Log-transformed pdfs for example BB7 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB7}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB7.pdf}
\caption[Example SBB7 copulas]{Log-transformed pdfs for example SBB7 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB7}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB8.pdf}
\caption[Example BB8 copulas]{Log-transformed pdfs for example BB8 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB8}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB8.pdf}
\caption[Example SBB8 copulas]{Log-transformed pdfs for example SBB8 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB8}}
\end{center}
\end{figure}

@@

\noindent We here define
\begin{equation}\label{eq.Cor}
   \cor_{l_b,u_b}(u,v) = \frac{\sum 
   (u_i-\mean(u)) (v_i-\mean(v))}{(n-1)\sqrt{\var(u)\var(v)}}.
\end{equation}
\noindent Here means and sample variances are computed using all $n$ data 
points, but the
sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
$u_i+v_i < 2u_b$. 
The sum of 
$\cor_{0,0.5}(u,v)$ and $\cor_{0.5,1}(u,v)$ (or some other choice of 
$\cor_{l_{b_k},u_{b_k}}(u,v)$ for bounds $l_{b_k},u_{b_k}$ that partition the
interval $(0,1)$) equals the standard Spearman correlation.

We also define $\Ps_{l_b,u_b}$, using the bounds
$u+v=2l_b$ and $u+v=2u_b$ defined in the main text, as follows. For a distance $h$, we define
$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
and a distance less than $h$ from the line $v=u$, divided by the 
total number of points within the bounds. This function is
defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
is half the longer of the two segments obtained by intersecting the
bounds with the unit square. It is easy to see $S(0)=0$ and 
$S(h_{\text{max}})=1$. We define $S_i(h)$, also for $h$ going from $0$
to $h_{\text{max}}$, to be the area within the bounds and the unit square
and within distance $h$ of the line $v=u$, divided by the area
within the bounds and the unit square. This is the expected value of
$S(h)$ for independent data. We then define
\begin{equation}\label{eq.P}
   P_{l_b,u_b} = \int_0^{h_{\text{max}}} ( S(h) - S_i(h) ) dh.
\end{equation}
\noindent As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
indicate stronger positive association
between $u$ and $v$ in the region given by the bounds.
We use both statistics 
because $\cor_{l_{b_k},u_{b_k}}(u,v)$ is more familiar, but 
in some instances $\Ps_{l_b,u_b}$ appears to have more power to
reveal tail dependence.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
\caption[For definitions of nonparametric statistics]{Schematic diagram supporting definitions of nonparametric tail dependence statistics. The blue and red lines are, respectively, $u+v=2l_b$ and $u+v=2u_b$.\label{fig_stat_image}}
\end{center}
\end{figure}

>>>>

<!--Now the copula figures that were not presented in the main text-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvClayton.pdf}
\caption[Example SC copulas]{Log-transformed pdfs (A-E) and samples (F-J) 
from example SC 
copulas. S is Spearman correlation; p is the value of the parameter for the SC 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively. The parameter range for the family is $p \in (0,\infty)$, upper-tail dependence is $2^{-1/p}$ and lower-tail
dependence is $0$.\label{figped_SC}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Gumbel.pdf}
\caption[Example G copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example G copulas. S is 
Spearman correlation; $p$ is the value of the parameter for the G family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter range for the family is $p \in [1,\infty)$, lower-tail dependence is $0$ and upper-tail
dependence is $2-2^{1/p}$.\label{figped_G}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvGumbel.pdf}
\caption[Example SG copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SG 
copulas. S is Spearman correlation; $p$ is the value of the parameter for the SG 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively. The parameter range for the family is $p \in [1,\infty)$, upper-tail dependence is $0$ and lower-tail
dependence is $2-2^{1/p}$.\label{figped_SG}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Joe.pdf}
\caption[Example J copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example J copulas. K is Kendall correlation; $p$ is the value of the parameter for the J family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter range for the family is $p \in [1,\infty)$, lower-tail dependence is $0$ and upper-tail
dependence is $2-2^{1/p}$.\label{figped_J}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvJoe.pdf}
\caption[Example SJ copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SJ 
copulas. K is Spearman correlation; $p$ is the value of the parameter for the SJ 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively. The parameter range for the family is $p \in [1,\infty)$, upper-tail dependence is $0$ and lower-tail
dependence is $2-2^{1/p}$.\label{figped_SJ}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Frank.pdf}
\caption[Example F copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example F copulas. S is Kendall correlation; $p$ is the value of the parameter for the F family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter range for the family is $p \in (0,\infty)$, lower-tail dependence is $0$ and upper-tail
dependence is $0$.\label{figped_F}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB1.pdf}
\caption[Example SBB1 copulas]{Log-transformed pdfs for example SBB1 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter ranges for the family are $p_1 \in (0,\infty)$
and $p_2 \in [1,\infty)$, lower-tail dependence is $2-2^{1/p_2}$ and upper-tail
dependence is $2^{-1/(p_1 p_2)}$.\label{figped_SBB1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB6.pdf}
\caption[Example BB6 copulas]{Log-transformed pdfs for example BB6 copulas. K is Kendall correlation; $p_1$ and $p_2$
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
and $p_2 \in [1,\infty)$, lower-tail dependence is $0$ and upper-tail
dependence is $2-2^{1/(p_1 p_2)}$. \label{figped_BB6}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB6.pdf}
\caption[Example SBB6 copulas]{Log-transformed pdfs for example SBB6 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
and $p_2 \in [1,\infty)$, upper-tail dependence is $0$ and lower-tail
dependence is $2-2^{1/(p_1 p_2)}$.\label{figped_SBB6}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB7.pdf}
\caption[Example BB7 copulas]{Log-transformed pdfs for example BB7 copulas. K is Kendall correlation; $p_1$ and $p_2$
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
and $p_2 \in (0,\infty)$, lower-tail dependence is $2^{-1/p_2}$ and upper-tail
dependence is $2-2^{1/p_1}$.\label{figped_BB7}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB7.pdf}
\caption[Example SBB7 copulas]{Log-transformed pdfs for example SBB7 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
and $p_2 \in (0,\infty)$, upper-tail dependence is $2^{-1/p_2}$ and lower-tail
dependence is $2-2^{1/p_1}$.\label{figped_SBB7}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB8.pdf}
\caption[Example BB8 copulas]{Log-transformed pdfs for example BB8 copulas. K is Kendall correlation; $p_1$ and $p_2$
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
and $p_2 \in (0,1]$, lower-tail dependence is $0$ and upper-tail
dependence is $2-2^{1/p_1}$ of $p_2=1$ and $0$ otherwise.\label{figped_BB8}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB8.pdf}
\caption[Example SBB8 copulas]{Log-transformed pdfs for example SBB8 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
and $p_2 \in (0,1]$, upper-tail dependence is $0$ and lower-tail
dependence is $2-2^{1/p_1}$ of $p_2=1$ and $0$ otherwise.\label{figped_SBB8}}
\end{center}
\end{figure}

@@

\noindent We define $\Ps_{l_b,u_b}$, using the bounds
$u+v=2l_b$ and $u+v=2u_b$ defined in the main text, as follows. For a distance $h$, we define
$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
and a distance less than $h$ from the line $v=u$, divided by the 
total number of points within the bounds. This function is
defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
is half the longer of the two segments obtained by intersecting the
bounds with the unit square. It is easy to see $S(0)=0$ and 
$S(h_{\text{max}})=1$. We define $S_i(h)$, also for $h$ going from $0$
to $h_{\text{max}}$, to be the area within the bounds and the unit square
and within distance $h$ of the line $v=u$, divided by the area
within the bounds and the unit square. This is the expected value of
$S(h)$ for independent data. We then define
\begin{equation}\label{eq.P}
   P_{l_b,u_b} = \int_0^{h_{\text{max}}} ( S(h) - S_i(h) ) dh.
\end{equation}
\noindent As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
indicate stronger positive association
between $u$ and $v$ in the region given by the bounds.
We use both statistics 
because $\cor_{l_{b_k},u_{b_k}}(u,v)$ is more familiar, but 
in some instances $\Ps_{l_b,u_b}$ appears to have more power to
reveal tail dependence.

<!--Decided we do not need this fig.
\begin{figure}[!h]
\begin{center}
\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
\caption[For definitions of nonparametric statistics]{Schematic diagram supporting definitions of nonparametric tail dependence statistics. The blue and red lines are, respectively, $u+v=2l_b$ and $u+v=2u_b$.\label{fig_stat_image}}
\end{center}
\end{figure}-->
